---
title: "s07 Exercises: Tibble Joins"
output: 
  html_document:
    keep_md: true
    theme: paper
---

## Requirements

You will need data from Joey Bernhardt's `singer` R package for this exercise. 

You can download the singer data from the `USF-Psych-DlassroomataSci/C` repo:

``` 
songs <- read_csv("https://github.com/USF-Psych-DataSci/Classroom/raw/master/data/singer/songs.csv", na = c("NA"))
locations <- read_csv("https://github.com/USF-Psych-DataSci/Classroom/raw/master/data/singer/loc.csv")
```

If you want, you could instead install the `singer` package itself. To do that,
you'll need to install `devtools`. Running this code in your console should do 
the trick:
```{r}
# update.packages(ask = FALSE)
```

``` {r, echo = FALSE}
# install.packages("devtools")
# devtools::install_github("JoeyBernhardt/singer")
```

Load required packages:

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
# library(singer)
knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, fig.align = "center")
```

<!-- The following chunk allows errors when knitting -->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

## Exercise 1: `singer`

The package `singer` comes with two smallish data frames about songs. 
Let's take a look at them (after minor modifications by renaming and shuffling):

```{r}
(time <- as_tibble(songs) %>% 
   rename(song = title))
```

```{r}
(album <- as_tibble(locations) %>% 
   select(title, everything()) %>% 
   rename(album = release,
          song  = title))
```


1. We really care about the songs in `time`. But, for which of those songs do we 
   know its corresponding album?

```{r}
time %>% 
  inner_join(album, by = c("song", "artist_name"))
```

2. Go ahead and add the corresponding albums to the `time` tibble, being sure to 
   preserve rows even if album info is not readily available.

```{r}
time %>% 
  left_join(album, by = c("song", "artist_name"))
```

3. Which songs do we have "year", but not album info?

```{r}
time %>% 
  anti_join(album, by = c("song", "artist_name"))
```

4. Which artists are in `time`, but not in `album`?

```{r}
time %>% 
  anti_join(album, by = "artist_name")
```


5. You've come across these two tibbles, and just wish all the info was 
   available in one tibble. What would you do?

```{r}
time %>% 
  full_join(album, by = c("song", "artist_name"))

# or

time %>% 
  full_join(select(album, -artist_name), by = "song")


```


## Exercise 2: LOTR

Load in three tibbles of data on the Lord of the Rings:

```{r}
fell <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv")
ttow <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv")
retk <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv")
```

1. Combine these into a single tibble.

```{r}
bind_rows(fell, ttow, retk)
```

2. Which races are present in "The Fellowship of the Ring" (`fell`), but not in 
   any of the other ones?

```{r}
fell %>% 
  anti_join(ttow, by = "Race") %>% 
  anti_join(retk, by = "Race")

# or

setdiff(fell$race, ttow$race)
```


## Exercise 3: Set Operations

Let's use three set functions: `intersect`, `union` and `setdiff`. We'll work 
with two toy tibbles named `y` and `z`, similar to Data Wrangling Cheatsheet

```{r}
(y <-  tibble(x1 = LETTERS[1:3], x2 = 1:3))

(z <- tibble(x1 = c("B", "C", "D"), x2 = 2:4))
```

1. Rows that appear in both `y` and `z`

```{r}
intersect(y, z)
```

2. You collected the data in `y` on Day 1, and `z` in Day 2. 
   Make a data set to reflect that.

```{r}
union(
  mutate(y, day = "Day 1"),
  mutate(z, day = "Day 2")
)
```

3. The rows contained in `z` are bad! Remove those rows from `y`.

```{r}
setdiff(y, z)
```


## Exercise 4: Flight Extra
```{r}
# install.packages("nycflights13")
library(nycflights13)
```

```{r}
flights %>% 
  left_join(weather, by = c("year", "month", "day", "origin", "hour", "time_hour")) %>% 
  left_join(planes, by = "tailnum") %>% 
  left_join(airports, c("dest" = "faa")) %>% 
  left_join(airlines, b = c("carrier" = "name")) %>% 
  select(flights, everything())
```

